Nondeterministic Bigraphical Reactive Systems for Markov Decision Processes\thanks{Supported by organization x.}

### 2018-08-11 paper.tex
Describing Jupyter

\documentclass[runningheads]{llncs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{wasysym}
\usepackage[lighttt]{lmodern}
\usepackage{siunitx}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\begin{definition}[concrete link graph]
  A \emph{concrete link graph}
  \[ F = (V_F, E_F, \ctrl_F, \link_F) : X \to Y \]
  is a quadruple having an inner face $X$ and an outer face $Y$, both finite
  subsets of $\mathcal{X}$, called respectivelly the \emph{inner} and
  \emph{outer names} of the link graph. $F$ has finite sets $V_F \subset
  \mathcal{V}$ of nodes and $E_F \subset \mathcal{E}$ of edges, a control map
  $\ctrl_F : V_F \to \mathcal{K}$ and a \emph{link map}
  \[ \link_F : X \uplus P_F \to E_F \uplus Y, \]
  where $P_F \coloneqq \{ (v, i) \mid i \in \ar(\ctrl_F(v)) \}$ is the set of
  ports of $F$. Thus $(v, i)$ is the $i$th port of node $v$. We shall call $X
  \uplus P_F$ the \emph{points} of $F$, and $E_F \uplus Y$ its \emph{links}.
\end{definition}

@@

\begin{definition}[nondeterministic reaction rule]
  Let $A$ be a set of actions. A \emph{nondeterministic reaction rule}
  $\mathsf{R}$ is a tuple $(R, R', a, p)$, where $(R, R', p)$ is a probabilistic
  reaction rule, and $a \in A$ is an action. We also define a \emph{reaction
    reward function} $r : A \to \mathbb{R}$ that assigns a reward/cost to each
  action.
\end{definition}

@@

TODO: introduce BigraphER \cite{DBLP:conf/cav/SevegnaniC16}

@@

The MDP can be represented as an NBRS with BigraphER code in
Listing~\ref{example_listing}. Furthermore, the BigraphER's visualisation of the
(automatically generated) transition system is in Fig.~\ref{example_ts}.

@@

Visually, each state is represented by a white ellipse (the starting state is
highlighted with a thicker border), and each action (per state) is a grey
ellipse. Text inside state ellipses lists all predicates that are satisfied by
that state. Edges from actions to states are labelled with both the probability
and the name of the reaction rule. Transitions from either a fully generated
transition system or a simulation, labels from predicates, and state/transition
rewards can be exported to the various plain text formats of a probabilistic
model checker PRISM \cite{DBLP:conf/cav/KwiatkowskaNP11} for further
quantitative analysis.

TODO: introduce jupyter (with more/new screenshots)

\begin{figure}
  \centering
  \includegraphics{images/interface.jpg}
  \caption{The BigraphER Jupyter interface with syntax highlighting.}
  \label{interface}
\end{figure}

@@

TODO: explain the syntax of the language

@@

In order to impement nesting, a \texttt{Cell} contains a \texttt{Node}, which
can have any number of \texttt{Cell}s of its own. Since after entering a room,
you end up in a specific place inside the room (next to the door), we also need
to link up the \texttt{Node} with one of the \texttt{Cell}s it contains (see
Fig.~\ref{agent2_nesting}). Both \texttt{Cell}s and \texttt{Node}s support
having a single link for exactly this purpose. Thus, most of the \texttt{Cell}
ports are not linked to anything, which is implemented via closures (e.g.
\texttt{/x Cell\{x\}}). The initial map in Listing~\ref{agent2_initial} has a
corridor of three cells, with the agent on the west-most cell, and a door to a
room in the middle cell. One moves in and out of a room via a pair of actions
with reaction rules pictured in Fig.~\ref{go_in_out}.

>>>>

\documentclass[runningheads]{llncs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{color}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{wasysym}
\usepackage[lighttt]{lmodern}
\usepackage{siunitx}
\usepackage{hyperref}
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
 \renewcommand\UrlFont{\color{blue}\rmfamily}

@@

\begin{definition}[concrete link graph]
  A \emph{concrete link graph}
  \[ F = (V_F, E_F, \ctrl_F, \link_F) : X \to Y \]
  is a quadruple having an inner face $X$ and an outer face $Y$, both finite
  subsets of $\mathcal{X}$, called respectively the \emph{inner} and
  \emph{outer names} of the link graph. $F$ has finite sets $V_F \subset
  \mathcal{V}$ of nodes and $E_F \subset \mathcal{E}$ of edges, a control map
  $\ctrl_F : V_F \to \mathcal{K}$ and a \emph{link map}
  \[ \link_F : X \uplus P_F \to E_F \uplus Y, \]
  where $P_F \coloneqq \{ (v, i) \mid i \in \ar(\ctrl_F(v)) \}$ is the set of
  ports of $F$. Thus $(v, i)$ is the $i$th port of node $v$. We shall call $X
  \uplus P_F$ the \emph{points} of $F$, and $E_F \uplus Y$ its \emph{links}.
\end{definition}

@@

\begin{definition}[nondeterministic reaction rule]
  Let $A$ be a set of actions. A \emph{nondeterministic reaction rule}
  $\mathsf{R}$ is a tuple $(R, R', a, p)$, where $(R, R', p)$ is a probabilistic
  reaction rule, and $a \in A$ is an action. We also define a \emph{reaction
    reward function} $r : A \to \mathbb{R}$ that assigns a reward or cost to
  each action.
\end{definition}

@@

% TODO: introduce BigraphER \cite{DBLP:conf/cav/SevegnaniC16} and its syntax

@@

The MDP can be represented as an NBRS with BigraphER code in
Listing~\ref{example_listing}. BigraphER can construct and visualise the full
transition system, as in Fig.~\ref{example_ts}.

@@

Visually, each state is represented by a white ellipse (the starting state is
highlighted with a thicker border), and each action (per state) is a grey
ellipse. The labels inside state ellipses list all predicates that are
satisfied by that state. Edges from actions to states are labelled with both the
probability and the name of the reaction rule. Transitions from either a fully
generated transition system or a simulation, labels from predicates, and
state/transition rewards can be exported to the various plain text formats of a
probabilistic model checker PRISM \cite{DBLP:conf/cav/KwiatkowskaNP11} for
further quantitative analysis.

\subsection{Front end}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{images/jupyter.jpg}
  \caption{The BigraphER Jupyter notebook interface with syntax highlighting.}
  \label{interface}
\end{figure}

We introduce a new front end to BigraphER based on Jupyter notebooks
\cite{DBLP:conf/elpub/KluyverRPGBFKHG16}. The BigraphER kernel is an extension
to the OCaml kernel\footnote{\url{https://github.com/akabe/ocaml-jupyter}} and
is available on
GitHub\footnote{\url{https://github.com/dilkas/bigrapher-jupyter}}. It supports
a similar workflow to that of Jupyter notebooks for other languages, allowing
the user to divide the code into multiple \emph{cells} and have the definitions
persist in a \emph{buffer} to be used later. By default, the output of each cell
displays the graphical representation of all bigraphs and reaction rules defined
in that cell (see Fig. \ref{interface} for an example). The generated images are
stored in a local directory \texttt{jupyter-images/}. The kernel also supports a
number of \emph{magics}, i.e., custom commands that can be used at the top of a
cell:

\begin{description}
  \item[\texttt{\%states}] produces a state transition diagram (similar to the
    one in Fig. \ref{example_ts}). Mousing over a state shows what that
    state looks like as a bigraph.
  \item[\texttt{\%simulate n}] runs a simulation for \texttt{n} steps, when used
    on non-stochastic models. For stochastic BRSs
    \cite{DBLP:journals/entcs/KrivineMT08}, \texttt{n} is interpreted as a
    floating-point number, representing maximum simulation time. The produced
    diagram is a single path through the transition system with the same
    mouse-over functionality.
  \item[\texttt{\%output}] displays the full output produced by BigraphER. By
    default, only errors are displayed in the notebook.
  \item[\texttt{\%ocaml}] allows full backward compatibility with the OCaml
    kernel, including auto-complete and integrated documentation features that
    have been extended to the BigraphER OCaml API as well as plotting support
    with Archimedes\footnote{\url{http://archimedes.forge.ocamlcore.org/}}.
  \item[\texttt{\%clear}] clears the buffer before interpreting the code from
    the current cell.
\end{description}

@@

@@

In order to implement nesting, a \texttt{Cell} contains a \texttt{Node}, which
can have any number of \texttt{Cell}s of its own. Since after entering a room,
you end up in a specific place inside the room (next to the door), we also need
to link up the \texttt{Node} with one of the \texttt{Cell}s it contains (see
Fig.~\ref{agent2_nesting}). Both \texttt{Cell}s and \texttt{Node}s support
having a single link for exactly this purpose. Thus, most of the \texttt{Cell}
ports are not linked to anything, which is implemented via closures (e.g.
\texttt{/x Cell\{x\}}). The initial map in Listing~\ref{agent2_initial} has a
corridor of three cells, with the agent on the west-most cell, and a door to a
room in the middle cell. One moves in and out of a room via a pair of actions
with reaction rules pictured in Fig.~\ref{go_in_out}.
